The Assembly of Black Hole Mass and Luminosity Functions of High-redshift Quasars\\via Multiple Accretion Episodes

### 2022-12-27 ms.tex
review1225; k revised version

To determine the initial conditions of the early BH assembly,
the formation pathway of seed BHs has been extensively investigated by numerical simulations and semi-analytical studies
\citep{2006MNRAS.370..289B,2009MNRAS.396..343R,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T,2022arXiv220505717B}.
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
and (iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the formation of heavy seeds
accelerates in rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and frequency of halo mergers.
The mass distribution of seed BHs is expected to be substantially top-heavy, extending the upper mass to $\gtrsim 10^{4-5}~\Msun$
\citep{2021ApJ...917...60L, 2022arXiv220614459T}.
\red{
We note that metal enrichment of the pristine halos from internal and external star formation activities also regulates the IMF.
Further studies accounting for this effect are essential to improve our work.
}
%As a result, \citet{2021ApJ...917...60L} predicted a top-heavy mass distribution of seed BHs ranging from $10^2-10^5~\Msun$.

@@

We consider seed BH formation in the main progenitors of quasar host galaxies, following a semi-analytical model established by \citet{2021ApJ...917...60L}.
In the highly-biased, overdense regions of the universe, those progenitor halos are likely irradiated by intense H$_2$-photodissociating radiation 
from nearby star-forming galaxies and heat the interior gas by successive mergers. 
The two effects counteracting H$_2$ formation and cooling prevent gas collapse and delay prior star formation \citep[e.g.,][]{2014MNRAS.445..107V,2019Natur.566...85W}.
Under such peculiar circumstances, massive clouds collapse to the halo centers at high accretion rates and leave massive stars (presumably seed BHs) behind.
Our model takes into account various key physical processes on star formation such as halo merger heating, radiative cooling, and
(photo-)chemical reaction networks.
The time evolution of the gas dynamics and thermal states is calculated in a self-consistent way with the halo assembly history.
\red{
We note that the H$_2$-photodissociating radiation is calculated following \citet{2014MNRAS.442.2036D},
measuring the flux at a free-fall timescale after single star bursts in nearby halos. 
This treatment underestimates the photodissociating flux, since the assumption of old stellar population provides less UV photons than smoothly-proceeding star formation
\citep{2021MNRAS.503.5046L}.
Therefore, our scenario is conservative in estimating the enhanced seed BH mass.
}

In addition, we consider baryonic streaming motion relative to DM produced in the epoch of cosmic recombination at $z_{\rm rec} \simeq 1100$.
This effect also delays gas collapse and star formation in protogalaxies through injection of kinetic energy into gas
\citep[e.g.,][]{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2017Sci...357.1375H,2019MNRAS.484.3510S,2021ApJ...917...60L}.
The amplitude of the streaming velocity is set to $\vbsm = 0$ and $\vbsm=1~\sigma_{\rm bsm}$, where $\sigma_{\rm bsm}=30~{\rm km~s}^{-1} (1+z)/(1+z_{\rm rec})$
is the root-mean-square speed at a redshift of $z$ \citep{2010PhRvD..82h3520T}.
We calculate the effective sound speed of gas as $c_{\rm eff} = \{c_{\rm s}^2+v_{\rm tur}^2/3+\left(\alpha_0 \vbsm \right)^2 \}^{1/2}$ at the center of each progenitor halo,
where $c_{\rm s}$ is the thermal sound speed, $v_{\rm tur}$ is the turbulent velocity of gas (i.e., the specific kinetic energy accumulated through halo mergers), and
the coefficient is set to $\alpha_0=1$ \footnote[1]{The value of $\alpha_0=1.0$ is motivated by cosmological simulations of massive primordial stars
under fast streaming velocities \citep{2017Sci...357.1375H,2019MNRAS.484.3510S}.
Note that our previous study in \citet{2021ApJ...917...60L} adopted $\alpha_0=4.7$ \citep{2018ApJ...855...17H}.}.
When the gas core becomes gravitationally unstable, the evolution of the gas density profile is well described by 
the Penston-Larson self-similar solution \citep{1969MNRAS.144..425P,1969MNRAS.145..271L}. 
The mass accretion rate from the envelope can be written as $\dot{M}=c_{\rm eff}^3/G$,
where $G$ is the gravitational constant.

At the vicinity of an accreting protostar, an accretion disk forms due to angular momentum of the inflowing material from large scales
\red{
\citep{2010MNRAS.403...45S,2011Sci...331.1040C,2011ApJ...737...75G}.
}
The episodic nature of disk accretion decreases the time-averaged accretion rate onto the protostar as $\Mdot_{\star}=\eta \Mdot$, 
where we adopt $\eta=0.3$ as the conversion efficiency \red{due to the centrifugal barrier} \citep{2016MNRAS.459.1137S, 2022arXiv220614459T}. 
%\blue{
%  (Toyouchi+ find 0.4 actually.)
%}
The size evolution of an accreting protostar and the radiative feedback strength depend sensitively on whether
the accretion rate becomes higher than a critical value of $\dot{M}_{\mathrm{crit}}=0.04~\Msunyr$ 
\citep{2001ApJ...561L..55O,2013ApJ...778..178H,2013A&A...558A..59S,2015MNRAS.452..755S,2018MNRAS.474.2757H}.
For $\dot{M}_\star > \Mdot_\mathrm{crit}$, the stellar envelope inflates owing to rapid entropy injection through accreting matter,
keeping its surface temperature as low as $T_\mathrm{eff} \simeq 5000$ K. 
Since stellar UV radiation is hardly emitted from the cold surface, the accretion flow efficiently feeds the central star without being 
impeded by radiative feedback.
As a result, the stellar mass reaches $\sim 10^5~\Msun$ but is limited by the general-relativistic instability to
%
\begin{equation}
 M_{\star, \mathrm{GR}} \simeq\left[0.83~\log \left(\frac{\dot{M}_{\star}}{\Msunyr}\right)+2.48\right] \times 10^{5} ~\Msun,
\end{equation}
%
in the range of $\dot{M}_\star\gtrsim 0.1~\Msunyr$ \citep{2016PhRvD..94b1501S,2017ApJ...842L...6W,2019PASA...36...27W}.
On the other hand, for $\dot{M}_\star \leq \Mdot_\mathrm{crit}$, the protostar begins to contract by loosing its thermal energy via radiative diffusion
and increases its surface temperature to $T_\mathrm{eff} \sim 10^5$ K.
Hence, intense stellar UV radiation heats the disk surface and launches mass outflows, preventing mass supply from the disk to the star
\citep{2008ApJ...681..771M,2011Sci...334.1250H}. 
In this case, the stellar mass is determined by the balance between mass accretion and mass loss via photoevaporation \citep{2013ApJ...773..155T}.
The feedback-regulated mass can be approximated as
\footnote[2]{A recent study by \citet{2022arXiv220614459T} provides a more sophisticated prescription of the feedback-regulated stellar mass based on their radiation hydrodynamical simulations.}
%
\begin{equation}
M_{\star, \mathrm{fb}} \simeq 2.9 \times 10^{3} \Msun\left(\frac{\dot{M}_{\star}}{0.01 M_{\odot} \mathrm{yr}^{-1}}\right),
\end{equation}
%
see more details in \cite{2021ApJ...917...60L}.
For an intermediate accretion rate between $\Mdot_\mathrm{crit}$ and $0.1~\Msunyr$, 
we perform logarithmic interpolation in the $\Mdot_\star - M_\star$ plane to smoothly connect 
$M_{\star, \mathrm{fb}}$ and $M_{\star, \mathrm{GR}}$ at the boundaries.
At the end of the stellar lifetime, those massive primordial stars result in BHs without significant mass loss 
via stellar winds \citep{2003ApJ...591..288H,2015MNRAS.451.4086S}.
Therefore, we assume that the mass of a remant BH is equal to that of its stellar progenitor.

@@

\vspace{2mm}
\subsection{BH mass growth}
\label{sec:model}
With BH seeds planted, we study the evolution of the BH mass distribution until $z=6$. 
The growth model of each BH is characterized by a minimal number of free parameters, 
giving the accretion rate by
%
\begin{equation}
  \label{eq:mdot}
  \Mdot_\bullet = \lambda f(\Mbh) \Mdot_\mathrm{Edd} ,
\end{equation}
where $\lambda$ is the ratio of the quasar bolometric luminosity to its Eddington luminosity $L_{\rm Edd}$,
% $\lambda \equiv L/L_\mathrm{Edd}$ 
and $\Mdot_\mathrm{Edd} \equiv L_{\mathrm{Edd}}/\eta_0 c^2$ is the Eddington accretion rate with a radiative efficiency of $\eta_0=0.1$ \citep{1973A&A....24..337S},
which is consistent with the value obtained from the Soltan-Paczy{\'n}ski argument \citep[e.g.,][]{2002MNRAS.335..965Y,2010ApJ...725..388C}.
\red{
Here we take the mass accretion rate as the net BH growth rate,
omitting the subtraction of energy loss with radiation by a factor of $1-\eta_0=0.9$.
This treatment simplifies the functional form without significantly impacting the results.
}
Based on studies on the mass dependence of the radiative efficiency for AGNs at $z\lesssim$ 3 
\citep{2008MNRAS.390..561C,2012ApJ...749..187L,2014ApJ...786..104U}, 
we introduce a similar functional form of
%
\begin{equation}
\label{eq:f_M}
f(\Mbh) = \frac{2}{1+\left(\Mbh /M_{\bullet,c} \right)^\delta}, 
\end{equation}
%
where $M_{\bullet,c}=10^8~\Msun$ is adopted
\footnote[3]{
A series of radiation hydrodynamical simulations for BH accretion
suggest that the gas supplying rate from galactic scales into the nuclear scale is not high enough to sustain 
super-Eddington accretion when the BH mass is higher than the characteristic value \citep{2021ApJ...907...74T}.}.
With a positive value of $\delta$, the BH growth at the high mass end is suppressed,
while the growth speed for less masssive BHs is accelerated.
In the limit of $\delta \ll 1$, where $f(\Mbh) \simeq 1$ is nearly independent of $M_{\bullet,c}$,
the model in Eq.~(\ref{eq:f_M}) reproduces an exponential growth with an $e$-folding timescale of
$t_{\rm S} =  \Mbh/\Mdot_{\mathrm{Edd}} \approx 45$ Myr (the so-called Salpeter timescale; \citealt{1964ApJ...140..796S}).

@@

To characterize the episodic accretion patterns of individual BHs, 
we introduce a time duration of $\tlife$, during which a single value of $\lambda$ is assigned to an accreting BH
following the ERDF of Eq.~(\ref{eq:Pl}).
In this way, we give a different growth speed of a BH in every period with a duration of $\tlife$ until $z=6$ 
(the corresponding cosmic age is $t_{\rm H}\simeq 913$ Myr).
This treatment enables us to capture the nature of accretion bursts and prohibit an unrealistically 
long-lasting rapid growth phase of a BH.
This phenomenological method avoids numerous uncertainties in modeling of the galaxy assembly, 
gas feeding, and BH feedback processes in the quasar progenitor halos,
\red{
which are implemented as sub-grid models in cosmological simulations \citep{2017MNRAS.467.4243D,2018MNRAS.473.4003B,2019MNRAS.488.4004L}.
Interestingly, these studies found the growth of massive BH seeds toward SMBHs is characterized by patterns including quiescent stages and accretion bursts.
Instead of treating these ingredients explicitly in our semi-analytical model,} we extract 
an averaged but fundamental timescale that governs mutual correlation between the QLF and BHMF,
based on direct comparison to those distribution functions for the observed high-$z$ quasar population.

@@

Using the BHMF and its evolution, we construct the QLF, which can be directly probed by high-$z$ quasar observations.
For an accreting BH with a bolometric luminosity of $\Lbol=\lambda L_\mathrm{Edd}$, 
its rest-frame ultraviolot (UV) absolute magnitude of $\Muv$ at 1450 $\mathrm{\AA}$ is estimated as
%
\begin{equation}
  \label{eq:M1450}
  \Muv= -21.0-2.5 \log  \left(\frac{\Lbol}{10^{45}~\mathrm{erg~s}^{-1}} \right) ~[\rm{mag}],
\end{equation}
%
where a bolometric correction factor of $f^{\rm bol}_{1450}=4.4$ for the monochromatic UV band is adopted
\citep{2006ApJS..166..470R}.
We construct the QLF from the BHMF, combining this conversion factor and ERDF.
However, we note that the shape and normalization of the {\it intrinsic} QLF is not necessarily identical to the {\it observed} QLF
because quasar surveys with rest-frame UV-to-optical bands are not sensitive to the obscured quasar population,
i.e., the observed number of fainter quasars is largely reduced by the obscuration effect
\citep{2003ApJ...598..886U,2007A&A...463...79G,2008A&A...490..905H,2014ApJ...786..104U,2014MNRAS.437.3550M}.

@@

In this study, we adopt the obscuration fraction $\fobsc$ based on X-ray observations \citep{2014ApJ...786..104U}:
\red{
%
\begin{eqnarray}
\fobsc & = \text{min} & \left\{ \psi_{\rm max}, \right.  \\ \nonumber 
&   & \left. \text{max}\left[ \psi_{43.75}  - \beta (\log L_{\rm X} -43.75), \psi_{\rm min}     \right]\right\},
\end{eqnarray}
%
where the parameters are written as $\psi_{\rm max}=0.84$, $\psi_{\rm min}=0.2$, $\psi_{43.75}=0.73$, and $\beta=0.24$.
% %
% \begin{eqnarray}
% \fobsc & = \text{min} & \left\{ 0.84, \right.  \\ \nonumber 
% &   & \left. \text{max}\left[   0.73-0.24 \log \left(\frac{L_{\rm X}}{10^{43.75}L_\odot}\right),0.2      \right]\right\}
% \end{eqnarray}
% %
%
Here, we use the bolometric correction factor of the hard X-ray band:}
%
\begin{equation}
  f^{\rm bol}_{\rm X} \equiv \frac{\Lbol}{L_{\rm X}}=
  a\left[1+\left(\frac{\log \left(\Lbol / L_\odot\right)}{b}\right)^{c}\right],
\end{equation}
%
where $a=10.96$, $b = 11.93$, and $c = 17.79$ \citep[see Eq.~2 in][]{2020A&A...636A..73D}. 
% In our analysis, we neglect the contribution from Compton-thick AGNs with $N_\mathrm{H}>10^{24}$ cm$^{-2}$, 
% which are obscured even in the hard X-ray band and show possible dependence on quasar properties.
It is worth noting that the functional form of the obscuration fraction
is calibrated with X-ray selected ANGs at $z\lesssim 5$,
and the value for high-$z$ quasars is still under investigation.
%and the uncertainties would affect our analyses for fainter quasars.
We leave more comprehensive arguments about this effect for future work.

>>>>

To determine the initial conditions of the early BH assembly,
the formation pathway of seed BHs has been extensively investigated by numerical simulations and semi-analytical studies
\citep{2006MNRAS.370..289B,2009MNRAS.396..343R,2009ApJ...696.1798T,2012MNRAS.422.2051N,2014ApJ...781...60H,
2015MNRAS.448..568H,2018MNRAS.474.3825V,2021MNRAS.506..613S,2022arXiv220614459T,2022arXiv220505717B}.
Massive star formation episodes in the early universe is substantially modulated by external environmental effects such as
(i) H$_2$ photo-dissociating irradiation from nearby galaxies
\citep{2001ApJ...546..635O,2002ApJ...569..558O,2003Natur.425..812B,2010MNRAS.402.1249S,2014MNRAS.445..544S,2014MNRAS.445..107V,2016ApJ...832..134C},
(ii) supersonic baryonic streaming motion that delays gas collapse in halos
\citep{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2018ApJ...855...17H,2017MNRAS.471.4878S,2018MNRAS.479.4017I},
and (iii) dynamical heating caused by frequent halo mergers and turbulence injection led by cold accretion flows
\citep{2003ApJ...592..645Y,2010Natur.466.1082M,2015ApJ...810...51M,2019Natur.566...85W,2022Natur.607...48L}.
Those effects keep isothermal collapse ($T\sim 8000~{\rm K}$) of massive gas at high accretion rates of $\gtrsim 0.1~\Msunyr$
without vigorous fragmentation \citep{2014MNRAS.445L.109I,2015MNRAS.446.2380B,2016PASA...33...51L},
and allow supermassive stars with $\sim 10^5~\Msun$ (presumably seed BHs) to form at the halo centers 
\citep{2013ApJ...778..178H,2013A&A...558A..59S,2019PASA...36...27W,2022arXiv220614459T}.
Although the conditions required for heavy seed formation are thought to be too stringent to be realized in the typical regions of the early universe
\citep{2008MNRAS.391.1961D,2009ApJ...695.1430A,2015MNRAS.450.4350I},
recent studies by \citet{2021MNRAS.503.5046L} and \citet{2021ApJ...917...60L} found that the formation of heavy seeds
accelerates in rare, overdense regions of the universe, where non-linear galaxy clustering boosts the irradiation intensity
and frequency of halo mergers.
\blue{In such massive halos, intense cold streams through cosmic webs and supersonic turbulent motion promote the formation
of massive gas clouds that collapse into supermassive stars and seed BHs \citep{2009Natur.457..451D,2012MNRAS.422.2539I,2022Natur.607...48L}.}
The mass distribution of seed BHs is expected to be substantially top-heavy, extending the upper mass to $\gtrsim 10^{4-5}~\Msun$
\citep{2021ApJ...917...60L, 2022arXiv220614459T}.
\blue{
Metal enrichment of the progenitor halos due to internal and external star formation activity would regulate the formation of 
seed BHs and alter the characteristic of the mass distribution function (see also a quantitative argument on the low efficiency of enrichment in 
\citealt{2021ApJ...917...60L}).
Further studies with hydrodynamic simulations are required to improve our understanding of the efficiency of metal enrichment in quasar progenitor halos
\citep[e.g.,][]{2018MNRAS.475.4378C}.
}
%As a result, \citet{2021ApJ...917...60L} predicted a top-heavy mass distribution of seed BHs ranging from $10^2-10^5~\Msun$.

@@

We consider seed BH formation in the main progenitors of quasar host galaxies, following a semi-analytical model established by \citet{2021ApJ...917...60L}.
In the highly-biased, overdense regions of the universe, those progenitor halos are likely irradiated by intense H$_2$-photodissociating radiation 
from nearby star-forming galaxies and heat the interior gas by successive mergers. 
The two effects counteracting H$_2$ formation and cooling prevent gas collapse and delay prior star formation \citep[e.g.,][]{2014MNRAS.445..107V,2019Natur.566...85W}.
Under such peculiar circumstances, massive clouds collapse to the halo centers at high accretion rates and leave massive stars (presumably seed BHs) behind.
Our model takes into account various key physical processes on star formation such as halo merger heating, radiative cooling, and
(photo-)chemical reaction networks.
The time evolution of the gas dynamics and thermal states is calculated in a self-consistent way with the halo assembly history.
\blue{
We calculate the time-dependent H$_2$-photodissociating and H$^-$ photo-detachment radiation flux (${\rm H}_2+\gamma \rightarrow 2~{\rm H}$; 
${\rm H}^-+\gamma \rightarrow {\rm H}+{\rm e}^-$) from each star forming galaxy with a radiation temperature of $T_{\rm rad}=2\times 10^4~{\rm K}$ \citep{2014MNRAS.445..544S,2015MNRAS.450.4350I}.
Following the treatment adopted in \citet{2014MNRAS.442.2036D}, the flux is measured at one free-fall timescale of the individual source halos after their single star bursts.
This assumption underestimates the radiation flux because the radiation contribution from pre-existing old stellar populations is neglected
(see also \citealt{2021MNRAS.503.5046L}, where the star formation activity is assumed to proceed in a smooth way rather than single burst episodes).
Therefore, our model gives a conservative estimate of the formation efficiency of seed BHs.
}

In addition, we consider baryonic streaming motion relative to DM produced in the epoch of cosmic recombination at $z_{\rm rec} \simeq 1100$.
This effect also delays gas collapse and star formation in protogalaxies through injection of kinetic energy into gas
\citep[e.g.,][]{2012MNRAS.424.1335F, 2014MNRAS.439.1092T, 2017Sci...357.1375H,2019MNRAS.484.3510S,2021ApJ...917...60L}.
The amplitude of the streaming velocity is set to $\vbsm = 0$ and $\vbsm=1~\sigma_{\rm bsm}$, where $\sigma_{\rm bsm}=30~{\rm km~s}^{-1} (1+z)/(1+z_{\rm rec})$
is the root-mean-square speed at a redshift of $z$ \citep{2010PhRvD..82h3520T}.
We calculate the effective sound speed of gas as $c_{\rm eff} = \{c_{\rm s}^2+v_{\rm tur}^2/3+\left(\alpha_0 \vbsm \right)^2 \}^{1/2}$ at the center of each progenitor halo,
where $c_{\rm s}$ is the thermal sound speed, $v_{\rm tur}$ is the turbulent velocity of gas (i.e., the specific kinetic energy accumulated through halo mergers), and
the coefficient is set to $\alpha_0=1$ \footnote{The value of $\alpha_0=1.0$ is motivated by cosmological simulations of massive primordial stars
under fast streaming velocities \citep{2017Sci...357.1375H,2019MNRAS.484.3510S}.
Note that our previous study in \citet{2021ApJ...917...60L} adopted $\alpha_0=4.7$ \citep{2018ApJ...855...17H}.}.
When the gas core becomes gravitationally unstable, the evolution of the gas density profile is well described by 
the Penston-Larson self-similar solution \citep{1969MNRAS.144..425P,1969MNRAS.145..271L}. 
The mass accretion rate from the envelope can be written as $\dot{M}=c_{\rm eff}^3/G$,
where $G$ is the gravitational constant.

\blue{
At the vicinity of an accreting protostar, an accretion disk forms due to angular momentum of the inflowing material from large scales
\citep{2010MNRAS.403...45S,2011Sci...331.1040C,2011ApJ...737...75G}.}
The episodic nature of disk accretion decreases the time-averaged accretion rate onto the protostar as $\Mdot_{\star}=\eta \Mdot$, 
where we adopt $\eta=0.3$ as the conversion efficiency \blue{owing to angular momentum of the accreting flow} \citep{2016MNRAS.459.1137S, 2022arXiv220614459T}. 
%\blue{
%  (Toyouchi+ find 0.4 actually.)
%}
The size evolution of an accreting protostar and the radiative feedback strength depend sensitively on whether
the accretion rate becomes higher than a critical value of $\dot{M}_{\mathrm{crit}}=0.04~\Msunyr$ 
\citep{2001ApJ...561L..55O,2013ApJ...778..178H,2013A&A...558A..59S,2015MNRAS.452..755S,2018MNRAS.474.2757H}.
For $\dot{M}_\star > \Mdot_\mathrm{crit}$, the stellar envelope inflates owing to rapid entropy injection through accreting matter,
keeping its surface temperature as low as $T_\mathrm{eff} \simeq 5000$ K. 
Since stellar UV radiation is hardly emitted from the cold surface, the accretion flow efficiently feeds the central star without being 
impeded by radiative feedback.
As a result, the stellar mass reaches $\sim 10^5~\Msun$ but is limited by the general-relativistic instability to
%
\begin{equation}
 M_{\star, \mathrm{GR}} \simeq\left[0.83~\log \left(\frac{\dot{M}_{\star}}{\Msunyr}\right)+2.48\right] \times 10^{5} ~\Msun,
\end{equation}
%
in the range of $\dot{M}_\star\gtrsim 0.1~\Msunyr$ \citep{2016PhRvD..94b1501S,2017ApJ...842L...6W,2019PASA...36...27W}.
On the other hand, for $\dot{M}_\star \leq \Mdot_\mathrm{crit}$, the protostar begins to contract by loosing its thermal energy via radiative diffusion
and increases its surface temperature to $T_\mathrm{eff} \sim 10^5$ K.
Hence, intense stellar UV radiation heats the disk surface and launches mass outflows, preventing mass supply from the disk to the star
\citep{2008ApJ...681..771M,2011Sci...334.1250H}. 
In this case, the stellar mass is determined by the balance between mass accretion and mass loss via photoevaporation \citep{2013ApJ...773..155T}.
The feedback-regulated mass can be approximated as
\footnote{A recent study by \citet{2022arXiv220614459T} provides a more sophisticated prescription of the feedback-regulated stellar mass based on their radiation hydrodynamical simulations.}
%
\begin{equation}
M_{\star, \mathrm{fb}} \simeq 2.9 \times 10^{3} \Msun\left(\frac{\dot{M}_{\star}}{0.01 M_{\odot} \mathrm{yr}^{-1}}\right),
\end{equation}
%
see more details in \cite{2021ApJ...917...60L}.
For an intermediate accretion rate between $\Mdot_\mathrm{crit}$ and $0.1~\Msunyr$, 
we perform logarithmic interpolation in the $\Mdot_\star - M_\star$ plane to smoothly connect 
$M_{\star, \mathrm{fb}}$ and $M_{\star, \mathrm{GR}}$ at the boundaries.
At the end of the stellar lifetime, those massive primordial stars result in BHs without significant mass loss 
via stellar winds \citep{2003ApJ...591..288H,2015MNRAS.451.4086S}.
\blue{One important caveat is the effect of stellar rotation on the structure evolution.
Namely, the mass loss rate of a rotating massive star via winds would be enhanced by the centrifugal force on the surface
or suppressed by efficient mixing of the interior structure by meridional circulation \citep[see][references therein]{2008A&A...489..685E,2012A&A...542A.113Y}.
Although those effects in the high mass regime are poorly understood yet, full general-relativistic simulations of the gravitational collapse of a rotating supermassive star 
show that most of the stellar mass is eventually swallowed by the newly born BH, ejecting only $\sim 10\%$ of the mass \citep{2002ApJ...572L..39S,2016PhRvD..94b1501S}.}
Therefore, we assume that the mass of a remnant BH is equal to that of its stellar progenitor.
%\footnote{
%\blue{Stellar rotation is expected to induce various effects on the structure evolution; namely, the mass loss rate via winds would be enhanced by the centrifugal force on the surface
%or suppressed by efficient mixing of the interior structure by meridional circulation \citep[see][references therein]{2008A&A...489..685E,2012A&A...542A.113Y}.
%Although those effects in the high mass regime are poorly understood yet, full general-relativistic simulations of the gravitational collapse of a rotating supermassive star 
%found that most of the stellar mass is eventually swallowed by the newly born BH, ejecting only $\sim 10\%$ of the mass \citep{2002ApJ...572L..39S,2016PhRvD..94b1501S}.}}.

@@

\vspace{2mm}
\subsection{BH mass growth}
\label{sec:model}
With BH seeds planted, we study the evolution of the BH mass distribution until $z=6$. 
The growth model of each BH is characterized by a minimal number of free parameters, 
giving the accretion rate by
%
\begin{equation}
  \label{eq:mdot}
  \Mdot_\bullet = \lambda f(\Mbh) \Mdot_\mathrm{Edd} ,
\end{equation}
where $\lambda$ is the ratio of the quasar bolometric luminosity to its Eddington luminosity $L_{\rm Edd}$,
% $\lambda \equiv L/L_\mathrm{Edd}$ 
and $\Mdot_\mathrm{Edd} \equiv L_{\mathrm{Edd}}/\eta_0 c^2$ is the Eddington accretion rate with a radiative efficiency of $\eta_0=0.1$ \citep{1973A&A....24..337S},
which is consistent with the value obtained from the Soltan-Paczy{\'n}ski argument \citep[e.g.,][]{2002MNRAS.335..965Y,2010ApJ...725..388C}.
Based on studies on the mass dependence of the radiative efficiency for AGNs at $z\lesssim$ 3 
\citep{2008MNRAS.390..561C,2012ApJ...749..187L,2014ApJ...786..104U}, 
we introduce a similar functional form of
%
\begin{equation}
\label{eq:f_M}
f(\Mbh) = \frac{2}{1+\left(\Mbh /M_{\bullet,c} \right)^\delta}, 
\end{equation}
%
where $M_{\bullet,c}=10^8~\Msun$ is adopted
\footnote{
A series of radiation hydrodynamical simulations for BH accretion
suggest that the gas supplying rate from galactic scales into the nuclear scale is not high enough to sustain 
super-Eddington accretion when the BH mass is higher than the characteristic value \citep{2021ApJ...907...74T}.}.
With a positive value of $\delta$, the BH growth at the high mass end is suppressed,
while the growth speed for less massive BHs is accelerated.
In the limit of $\delta \ll 1$, where $f(\Mbh) \simeq 1$ is nearly independent of $M_{\bullet,c}$,
the model in Eq.~(\ref{eq:f_M}) reproduces an exponential growth with an $e$-folding timescale of
$t_{\rm S} =  \Mbh/\Mdot_{\mathrm{Edd}} \approx 45$ Myr (the so-called Salpeter timescale; \citealt{1964ApJ...140..796S}).
\blue{
Note that we do not explicitly consider the energy loss with radiation in Eq.~(\ref{eq:mdot}), neglecting a factor of $1-\eta_0=0.9$.
This small deviation from unity can be absorbed by the uncertainty due to the functional form of $f(M_\bullet)$ and does not bring 
a significant impact on the results discussed below.
}

@@

To characterize the episodic accretion patterns of individual BHs, 
we introduce a time duration of $\tlife$, during which a single value of $\lambda$ is assigned to an accreting BH
following the ERDF of Eq.~(\ref{eq:Pl}).
In this way, we give a different growth speed of a BH in every period with a duration of $\tlife$ until $z=6$ 
(the corresponding cosmic age is $t_{\rm H}\simeq 913$ Myr).
This treatment enables us to capture the nature of accretion bursts and prohibit an unrealistically 
long-lasting rapid growth phase of a BH.
This phenomenological method avoids numerous uncertainties in modeling of the galaxy assembly, 
gas feeding, and BH feedback processes in the quasar progenitor halos,
\blue{
which are implemented as sub-grid models in previous cosmological simulations \citep{2017MNRAS.467.4243D,2018MNRAS.473.4003B,2019MNRAS.488.4004L}.
These studies found that the growth of massive BH seeds toward SMBHs is characterized by the intermittent patterns with quiescent and accretion phases.
Instead of treating these ingredients explicitly in our semi-analytical model,} we extract 
an averaged but fundamental timescale that governs mutual correlation between the QLF and BHMF,
based on direct comparison to those distribution functions for the observed high-$z$ quasar population.

@@

Using the BHMF and its evolution, we construct the QLF, which can be directly probed by high-$z$ quasar observations.
For an accreting BH with a bolometric luminosity of $\Lbol=\lambda L_\mathrm{Edd}$, 
its rest-frame ultraviolet (UV) absolute magnitude of $\Muv$ at 1450 $\mathrm{\AA}$ is estimated as
%
\begin{equation}
  \label{eq:M1450}
  \Muv= -21.0-2.5 \log  \left(\frac{\Lbol}{10^{45}~\mathrm{erg~s}^{-1}} \right) ~[\rm{mag}],
\end{equation}
%
where a bolometric correction factor of $f^{\rm bol}_{1450}=4.4$ for the monochromatic UV band is adopted
\citep{2006ApJS..166..470R}.
We construct the QLF from the BHMF, combining this conversion factor and ERDF.
However, we note that the shape and normalization of the {\it intrinsic} QLF is not necessarily identical to the {\it observed} QLF
because quasar surveys with rest-frame UV-to-optical bands are not sensitive to the obscured quasar population,
i.e., the observed number of fainter quasars is largely reduced by the obscuration effect
\citep{2003ApJ...598..886U,2007A&A...463...79G,2008A&A...490..905H,2014ApJ...786..104U,2014MNRAS.437.3550M}.

@@

In this study, we adopt the obscuration fraction $\fobsc$ based on X-ray observations \citep{2014ApJ...786..104U}:
\blue{
%
\begin{eqnarray}
%\fobsc & = \text{min} & \left\{ \psi_{\rm max}, \right.  \\ \nonumber 
%&   & \left. \text{max}\left[ \psi_{43.75}  - \beta (\log L_{\rm X} -43.75), \psi_{\rm min}     \right]\right\},
\fobsc && = \text{min}  \left[ \psi_{\rm max}, \psi \right], \\[5pt]
\psi &&= \text{max}\left[\psi_0  - \beta \log (L_{\rm X}/L_{\rm X,0}), \psi_{\rm min}     \right],
\end{eqnarray}
%
where the parameters are written as $\psi_{\rm max}=0.84$, $\psi_{\rm min}=0.2$, $\psi_{0}=0.73$, $\beta=0.24$, and $L_{\rm X,0}=10^{43.75}~{\rm erg~s}^{-1}$.
% %
% \begin{eqnarray}
% \fobsc & = \text{min} & \left\{ 0.84, \right.  \\ \nonumber 
% &   & \left. \text{max}\left[   0.73-0.24 \log \left(\frac{L_{\rm X}}{10^{43.75}L_\odot}\right),0.2      \right]\right\}
% \end{eqnarray}
% %
%
Here, we use the bolometric correction factor of the hard X-ray band:}
%
\begin{equation}
  f^{\rm bol}_{\rm X} \equiv \frac{\Lbol}{L_{\rm X}}=
  a\left[1+\left(\frac{\log \left(\Lbol / L_\odot\right)}{b}\right)^{c}\right],
\end{equation}
%
where $a=10.96$, $b = 11.93$, and $c = 17.79$ \citep[see Eq.~2 in][]{2020A&A...636A..73D}. 
% In our analysis, we neglect the contribution from Compton-thick AGNs with $N_\mathrm{H}>10^{24}$ cm$^{-2}$, 
% which are obscured even in the hard X-ray band and show possible dependence on quasar properties.
It is worth noting that the functional form of the obscuration fraction
is calibrated with X-ray selected ANGs at $z\lesssim 5$,
and the value for high-$z$ quasars is still under investigation.
%and the uncertainties would affect our analyses for fainter quasars.
We leave more comprehensive arguments about this effect for future work.
